---
title: "Usage"
date: "`r format(Sys.time(), '%d-%b-%Y')`"
output:
  html_document:
    highlight: kate
    self_contained: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(patchwork)
library(plummy)
```

## Import

Import fonts the first time.

```{r, eval=FALSE}
library(extrafont)

plummy::import_inter()
plummy::import_plex()
plummy::import_nunito()
plummy::import_barlow()

# check that font names look right
extrafont::fonts()
extrafont::fonttable()
```

Load fonts for use. Run once per session.

```{r, message=FALSE, warning=FALSE}
extrafont::loadfonts()
```

## Use theme

Let's create some sample plots.

```{r,fig.height=9.5,fig.width=9}
p1 <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) +
    geom_point() +
    labs(
        title = "Iris dataset",
        subtitle = "Comparison of iris species",
        x = "Sepal length",
        y = "Sepal width",
        caption = "Source: Edgar Anderson"
    )
p2 <- p1 + theme_inter()
p3 <- p1 + theme_plex()
p4 <- p1 + theme_nunito()

wrap_plots(p1, p2, p3, p4, nrow = 2, ncol = 2)
```

**Fig:** Comparison of default theme, inter, plex and nunito.

To set theme for all plots in a session, use

```{r,eval=FALSE}
library(ggplot2)
theme_set(theme_inter())
```

## Use palette

```{r,fig.height=9.5,fig.width=9}
p1 <- ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
    geom_point() +
    labs(
        title = "Iris dataset",
        subtitle = "Comparison of iris species",
        x = "Sepal length",
        y = "Sepal width",
        caption = "Source: Edgar Anderson"
    )+
    theme(legend.position = "top")
p2 <- p1 + scale_color_tableau() + theme_inter(grid = "Y", leg = "h")
p3 <- p1 + scale_color_tableau() + theme_plex(grid = "Y", leg = "h")
p4 <- p1 + scale_color_tableau() + theme_nunito(grid = "Y", leg = "h")

wrap_plots(p1, p2, p3, p4, nrow = 2, ncol = 2)
```

**Fig:** Custom color scale, legend in placed on the top and only horizontal gridlines are displayed.

```{r,fig.height=5,fig.width=12}
p2 <- p1 + scale_color_tableau() + theme_inter(axis = FALSE, grid = FALSE, leg = "h")
p3 <- p1 + scale_color_tableau() + theme_plex(axis = FALSE, grid = FALSE, leg = "h")
p4 <- p1 + scale_color_tableau() + theme_nunito(axis = FALSE, grid = FALSE, leg = "h")

wrap_plots(p2, p3, p4, nrow = 1)
```

**Fig:** Plain theme with no axis and no gridlines.

```{r,fig.height=9.5,fig.width=10}
p1 <- ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
    geom_point() +
    facet_grid(~Species)+
    labs(
        title = "Iris dataset",
        subtitle = "Comparison of iris species",
        x = "Sepal length",
        y = "Sepal width",
        caption = "Source: Edgar Anderson"
    )+
    theme(legend.position = "top")
p2 <- p1 + scale_color_tableau() + theme_inter(leg = "h")
p3 <- p1 + scale_color_tableau() + theme_plex(leg = "h")
p4 <- p1 + scale_color_tableau() + theme_nunito(leg = "h")

wrap_plots(p1, p2, p3, p4, nrow = 2, ncol = 2)
```

**Fig:** Example with facetting.

To set palette for all plots in a session, use

```{r,eval=FALSE}
library(ggplot2)
options(ggplot2.discrete.colour = palette_tableau()(20))
options(ggplot2.discrete.fill = palette_tableau()(20))
```

## Fonts

Available fonts in the package can be listed.

```{r}
list_fonts()
```

Below is a preview of fonts.

```{r,fig.height=8,fig.width=9,echo=FALSE}
dfr <- list_fonts()
dfr$variant <- factor(dfr$variant,levels=c("thin","light","regular","medium","semibold"))
dfr1 <- dfr[dfr$variant=="regular",]
dfr1$label <- paste(c(paste(LETTERS, collapse = ""), "ÅÖÄ\n", paste(letters, collapse = ""), "åöä\n1234567890", "!\"#¤%&/()=?+*~$£@{}[]"), collapse = "")

ggplot(dfr1) +
    geom_text(aes(x = variant, y = font, label = label,family = font), size = 7.5) +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank()
    )
```

Below is an overview of all fonts and it's variants.

```{r,fig.height=5,fig.width=10,echo=FALSE}
ggplot(dfr,aes(x=family,y=variant))+
    geom_text(aes(label=font, family=font),size=5)+
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.title = element_blank()
    )
```

It is also possible to cusstomise the themes avilable within the package or to use custom fonts by providing font names to the `family` arguments in `theme_custom()`.